---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yihang Cai
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(tidymodels)
library(palmerpenguins)
```


# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....



# Data {#sec-data}



# Model


## Model set-up

Define $y_i$ as the passing epa. Then $\beta_0$ is the interception and here the passing epa is normally distributed with a mean $\mu$ and a standard deviation $\sigma$, where the mean depends on three parameters $\beta_0$, $\beta_1$, $\beta_2$, and their attempts and sacks.   

\begin{align} 
y_i|\mu_i &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \alpha^{attempts}_i + \alpha^{sacks}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) 
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification



# Results
```{r}
#| message: false
#| echo: false
#| warning: false

linear_model <- readRDS("../models/linear_model.rds")
analysis_data <- read_parquet("../data/analysis_data/analysis_data.parquet")

analysis_data <- analysis_data %>%
  group_by(player_id) %>%
  mutate(
    passing_epa_rolling_avg = rollapply(passing_epa, width = 3, FUN = mean, partial = TRUE, align = 'right', fill = NA)
  ) %>%
  ungroup()

# Fill NA values in passing_epa_rolling_avg with the player's existing passing_epa values
analysis_data <- analysis_data %>%
  group_by(player_id) %>%
  mutate(passing_epa_rolling_avg = ifelse(is.na(passing_epa_rolling_avg), 0, passing_epa_rolling_avg)) %>%
  ungroup()

linear_preds <- predict(linear_model, nfl_testing_data)
results <- bind_cols(nfl_testing_data, linear_preds)
mae_results <- mae(results, truth = passing_epa, estimate = .pred)

print(results)
print(mae_results)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check



\newpage


# References


